### 2019-04-15 paper/paper.tex
various edits

To collect our data for package usage in public projects, 
we turned to the Google BigQuery \code{github\_repos} data 
set\footnote{https://bigquery.cloud.google.com/dataset/bigquery-public-dat:github\_repos}.
From this data, we skimmed all projects and pulled those 
written in JavaScript, resulting in an initial total of 9,017,221 
repositories. Simply being written in JavaScript is not enough to detect
\code{npm} usage, so to find repositories using \code{npm} packages
we looked for projects that contained a \code{package.json} 
file. After confirming that the formatting of the file matched the \code{npm}
standard, we are able to confidently say that any remaining packages are 
using \code{npm} packages. With this criteria, we then compiled a list 
of [100,000 for now] project 
repositories from GitHub, a volume that we felt is likely to be an 
appropriate representation of all such public projects.
Using the \code{GitHub} platform API and the list of valid repository names 
from before, we cloned the GitHub repository for each project. With
this, we were then able to determine the date and time of every 
commit of the project, as well as the list of included \code{npm} packages
at the time of commit, as listed in every historic version of 
\code{package.json}, providing us with the \code{npm} packages used by that repository
at any given time in the project's history. The results showed that on average a repository 
had 37.9 commits. Ultimately, of the 9,017,221 
public \code{GitHub} projects written in JavaScript and out of all of the repositories
using \code{npm} packages, we randomly selected 40,000 repositories as
a representative data set for the analysis.

@@

Created in 2009, \code{npm} has grown
rapidly in popularity and scope over the last ten years, and 
as of the original paper showed no signs of slowing down \cite{Wittern:2016}.
We investigate the state of the \code{npm} ecosystem
since September 1\textsuperscript{st} 2015, and look for any signs of deterioration in 
the health of the ecosystem. Periods of stagnating growth would suggest
that developer interest is waning, while steady activity would 
indicate that the ecosystem as a whole is healthy and will continue to
evolve. To search for these potential indicators in the \code{npm} package environment,
we investigate the number of packages created and updated over time, as well
as system-wide trends of dependencies within packages. 
In Figure \ref{npmGrowth},
we can observe that while \code{npm} is still growing, the speed at which new packages
are created has slowed from an exponential to a linear rate. Packages added and updated 
per month have gone from 10,087 and 23,100 respectively in September 2015 to 17,354 and 48,140 respectively as of April 2019.
Some other things of note include an observable spike in package creation and updates around March 2016, the month 
when developer Azer Ko\c{c}ulu removed his 273 packages (including the popular package 
\code{left-pad}) from \code{npm} and in doing so affected some packages such as 
\code{babel} and \code{atom}, and therefore a significant fraction of \emph{all} 
\code{Node.js} projects\footnote{https://blog.npmjs.org/post/141577284765/kik-left-pad-and-npm}.
There is also a significant spike in package updates observed in the month of January 2019, caused 
by some source currently unknown to the authors.
[ADD CONCAVITY DATA].

@@

\textbf{Takeaways.} We find that while the \code{npm} ecosystem
continues to grow, the rate of growth is no longer superlinear, 
a possible sign of the ecosystem beginning to reach some form of maturity \cite{Alves:2011}.
In the relationships between packages, we find that though the average number
of dependencies continues to grow, there has been a recent decline in the
rate of this growth. Accordingly, we have found that the number of packages with
zero dependencies has increased from 28.7\% in September 2015 to [DATA REQUIRED]\%
as of April 2019. The percentage of packages that are depended on has
also declined, albeit still exhibiting a power law relationship as observed
in prior software ecosystem research \cite{Wittern:2016, Myers:2003, Louridas:2008}.

@@

While these packages represent the most consistently popular
members of \code{npm}, we also evaluate the most popular packages
on a yearly basis. Figure \ref{ranksByYear} presents the top five packages
for each individual year from 2011 to 2019. As noted in the original paper,
the top packages per year are relatively stable from 2013 on \cite{Wittern:2016}.
One notable exception to this rule is the 2016 rise in popularity of \code{eslint},
a tool for improving the quality of \code{JavaScript} code\footnote{https://www.npmjs.com/package/eslint}.

@@

To better understand the evolution of package popularity,
we investigate the dynamics of the top rankings over time.
While figure \ref{ranksByYear} demonstrate that package popularity
appears to be relatively stable, we can examine the number of packages
entering (and therefore exiting) the top \code{n} packages per year in
order to obtain a better view of the ecosystem as a whole.
Table \ref{numEnteringTop} depicts the amount of newcomers to the top
\code{npm} ranks each year, further supporting the observation that
core \code{npm} package popularity is becoming more stable as the 
ecosystem matures.

@@

Finally, \code{npm} has been investigated informally. In
the \code{pagerank} package, web-based \code{PageRanks} are used to evaluate
package popularity\footnote{https://www.npmjs.com/package/pagerank}.
The project code{npm-by-the-numbers} also provides various statistics
about the state of \code{npm}, but is significantly out of date and 
in contrast to this work does not provide an evaluation of the ecosystem over time
or give any insight into application usage\footnote{https://github.com/ashleygwilliams/npm-by-the-numbers}.

@@

In this paper, we take a second look at the \code{npm} ecosystem,
replicating and extending the results of previous research by 
Wittern, Suter and Rajagopalan \cite{Wittern:2016}. Though
we observe that the growth of \code{npm} has slowed down, the ecosystem
continues to thrive, demonstrating signs of maturity rather than decline.

Contrary to the trend observed in the original paper, there appears to be
some effort to reduce the number of inter-package dependencies, and therefore the risks
associated with complex dependency trees. The core set of packages upon which
multiple other projects rely on has remained relatively constant in size, and
demonstrates a trend towards even further concentration. Accordingly, the set of packages
most depended on in the \code{npm} ecosystem is also trending towards stability, with
fewer and fewer packages entering the top ranks each year.

>>>>

To collect our data for package usage in public projects, 
we turned to the Google BigQuery \code{github\_repos} data 
set\footnote{https://bigquery.cloud.google.com/dataset/bigquery-public-dat:github\_repos}.
From this data, we skimmed all projects and pulled those 
written in JavaScript, resulting in an initial total of 9,017,221 
repositories. Simply being written in JavaScript is not enough to detect
\code{npm} usage, so to find repositories using \code{npm} packages
we looked for projects that contained a \code{package.json} 
file. After confirming that the formatting of the file matched the \code{npm}
standard, we are able to confidently say that any remaining packages are 
using \code{npm} packages. From this criteria, we compiled a list 
of applicable project 
repositories from GitHub.
Using the \code{GitHub} platform API and the list of valid repository names 
from before, we cloned the GitHub repository for a random subset of all projects. With
this, we were then able to determine the date and time of every 
commit of these projects, as well as the list of included \code{npm} packages
at the time of every commit, as listed in every historic version of 
\code{package.json}, providing us with the \code{npm} packages used by that repository
at any given time in the project's history. The results showed that on average a repository 
had 37.9 commits to its \code{package.json}. Ultimately, of the 9,017,221 
public \code{GitHub} projects written in JavaScript and out of all of the repositories
using \code{npm} packages, we selected 40,000 repositories as
a representative data set for the analysis.

@@

Created in 2009, \code{npm} has grown
rapidly in popularity and scope over the last ten years, and 
as of the original paper showed no signs of slowing down \cite{Wittern:2016}.
We investigate the state of the \code{npm} ecosystem
since September 1\textsuperscript{st} 2015, and look for any signs of deterioration in 
the health of the ecosystem. Periods of stagnating growth would suggest
that developer interest is waning, while steady activity would 
indicate that the ecosystem as a whole is healthy and will continue to
evolve. To search for these potential indicators in the \code{npm} package environment,
we investigate the number of packages created and updated over time, as well
as system-wide trends of dependencies within packages. 
In Figure \ref{npmGrowth},
we can observe that while \code{npm} is still growing, the speed at which new packages
are created has slowed from an exponential to a linear rate. Packages added and updated 
per month have gone from 10,087 and 23,100 respectively in September 2015 to 17,354 and 48,140 respectively as of April 2019.
Some other things of note include an observable spike in package creation and updates around March 2016, the month 
when developer Azer Ko\c{c}ulu removed his 273 packages (including the popular package 
\code{left-pad}) from \code{npm} and in doing so affected some packages such as 
\code{babel} and \code{atom}, and therefore a significant fraction of \emph{all} 
\code{Node.js} projects\footnote{https://blog.npmjs.org/post/141577284765/kik-left-pad-and-npm}.
There is also a significant spike in package updates observed in the month of January 2019, caused 
by some source currently unknown to the authors.
% [ADD CONCAVITY DATA].

@@

\textbf{Takeaways.} We find that while the \code{npm} ecosystem
continues to grow, the rate of growth is no longer superlinear, 
a possible sign of the ecosystem beginning to reach some form of maturity \cite{Alves:2011}.
In the relationships between packages, we find that though the average number
of dependencies continues to grow, there has been a recent decline in the
rate of this growth. Accordingly, we have found that the the percentage of packages with zero dependencies has 
increased from 19.5\% in September 2015 to 23.8\% as of April 2019. The ratio of packages that are depended on has
also declined, with the percentage of packages with
one or more dependants decreasing from 27.2\% in September 2015 to 20.8\%
as of April 2019. However, this still exhibits a power law relationship as observed
in prior software ecosystem research \cite{Wittern:2016, Myers:2003, Louridas:2008}.

@@

While these packages represent the most consistently popular
members of \code{npm}, we also evaluate the most popular packages
on a yearly basis. Figure \ref{ranksByYear} presents the top five packages
for each individual year from 2011 to 2019. As noted in the original paper,
the top packages per year are relatively stable from 2013 onwards \cite{Wittern:2016}.
One notable exception to this rule is the 2016 rise in popularity of \code{eslint},
a tool for improving the quality of \code{JavaScript} code\footnote{https://www.npmjs.com/package/eslint}.

@@

To better understand the evolution of package popularity,
we investigate the dynamics of the top rankings over time.
While Figure \ref{ranksByYear} demonstrates that package popularity
appears to be relatively stable, we can examine the number of packages
entering (and therefore exiting) the top \code{n} packages per year in
order to obtain a better view of the ecosystem as a whole.
Table \ref{numEnteringTop} depicts the amount of newcomers to the top
\code{npm} ranks each year, further supporting the observation that
core \code{npm} package popularity is becoming more stable as the 
ecosystem matures.

@@

Finally, \code{npm} has been investigated informally. In
the \code{pagerank} package, web-based PageRanks are used to evaluate
package popularity\footnote{https://www.npmjs.com/package/pagerank}.
The project \code{npm-by-the-numbers} also provides various statistics
about the state of \code{npm}, but is significantly out of date and 
in contrast to this work does not provide an evaluation of the ecosystem over time
or give any insight into application usage\footnote{https://github.com/ashleygwilliams/npm-by-the-numbers}.

@@

In this paper, we take a second look at the \code{npm} ecosystem,
replicating and extending the results of previous research by 
Wittern, Suter and Rajagopalan \cite{Wittern:2016}. Though
we observe that the growth of \code{npm} has slowed down, the ecosystem
continues to thrive, demonstrating signs of maturity rather than stagnation.

Contrary to the trend observed in the original paper, there now appears to be
some effort to reduce the number of inter-package dependencies, and therefore the risks
associated with complex dependency trees. The core set of packages upon which
multiple other projects rely on has remained relatively constant in size, and
demonstrates a trend towards even further concentration. Accordingly, the set of packages
most depended on in the \code{npm} ecosystem is also trending towards increasing stability, with
fewer and fewer packages entering the top ranks each year.
